---
title: "Lab4HugoOtto"
author: "Hugo Knape & Otto Moen"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Lab4HugoOtto}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Lab4HugoOtto)
```

# The linreg function
The Lab4HugoOtto package provides the linreg function which is used to fit linear models. The linreg function can take two arguments (formula, data) and returns a list with class linreg which include Regressions coefficients, the fitted values, the residuals, the degrees of freedom, the residual variance, the variance of the regression coefficients and the t-values for each coefficient. 

```{r}
linreg_obj <- linreg(Petal.Length~Species, iris)
```

# Methods for \code{linreg} objects

The package also provides several generic methods for objects of the class \code{linreg}, which can be used to analyze different aspects of the results from the linear regression. Here follows a short description of each method.

## \code{print.linreg}

The \code{print.linreg} method will return the estimated coefficients from the regression model along with their names.

```{r}
print(linreg_obj)
```

## \code{plot.linreg}

The \code{plot.linreg} method will return two plots using the \code{ggplot2} package. The two plots returned are the Residuals vs Fitted plot as well as the Scale-Location plot. 

```{r}
plot(linreg_obj)
```

## \code{residuals.linreg}

The \code{residuals.linreg} method will return the residuals from the regression model as a vector.

```{r}
resid(linreg_obj)
```


## \code{pred}

The \code{pred} method will return the predicted values from the regression model as a vector.

```{r}
pred(linreg_obj)
```

## \code{coef.linreg}

The \code{coef.linreg} method will return the coefficients from the regression model as a named vector.

```{r}
coef(linreg_obj)
```

# \code{theme_linkoping}

Finally the package provides a theme for plots created with the \code{ggplot2} package. This is a basic theme that controls the size of the plot title, the axis titles and the axis text. It also sets the colors of the plot title and axis title to a blue hue that matches the base color in the graphical manual of Linkoping University.     


